A formalization of semi-simplicial and semi-cubical sets

### 2023-01-06 paper/paper.tex
paper: more polish

\title{A mechanization of semi-simplicial and semi-cubical sets in the calculus of inductive constructions}
\author{
  \textcolor{gray80}{Hugo Herbelin} \\
  \itshape \textcolor{gray80}{Université Paris-Cité, Inria, CNRS, IRIF, Paris}
  \and
  \textcolor{gray80}{Ramkumar Ramachandra} \\
  \itshape \textcolor{gray80}{Université Paris-Cité, Inria, CNRS, IRIF, Paris}
}
\date{}

@@

\begin{document}
\maketitle
\begin{abstract}
  We present a parametricity-based construction of semi-simplicial sets and semi-cubical sets in indexed form, in the calculus of inductive constructions. The accompanying artifact is an axiom-less~\footnote{Technically speaking, we use some axioms related to functional extensionality, but the requirement arises only from our generalization to $\nu$-types; if $\nu$ is finite (say, unit type corresponding to augmented semi-simplicial types), this requirement vanishes.} Coq formalization that uses sophisticated rewriting techniques and pushes the boundaries of proof assistant technology.
\end{abstract}
\tableofcontents
\newpage

@@

The outline of the paper is as follows. First, we briefly describe semi-simplicial and semi-cubical sets, and our generalization to $\nu$-sets. We then proceed to explain our construction, with the help of diagrams. Before spelling out the details of our formalization in type-theoretic language, we have a section on reasoning in sets.

@@

An equivalent defintion of UIP is given by \textsf{UIP\_refl}:

@@

% Restriction for frame, layer, and filler
% Drop the unviverse letter
\renewcommandx{\restrf}[7][1,2,3,4,5,6,7]{\restr{frame}[][#2][#3][#4][#5][#6][#7]}
\renewcommandx{\restrl}[7][1,2,3,4,5,6,7]{\restr{layer}[][#2][#3][#4][#5][#6][#7]}
\renewcommandx{\restrc}[7][1,2,3,4,5,6,7]{\restr{filler}[][#2][#3][#4][#5][#6][#7]}

@@

We now formally present the construction in \emph{extensional} type theory; i.e. in a type theory with the following reflection rule, where $=$ is propositional equality in some type and $\equiv$ is definitional equality:

@@

@@

Since we have formalized the proof in the Coq proof assistant, which works in intensional type theory, we now present the tables in \emph{intensional} type theory, leaving implicit some details, which we will explain shortly.

@@

\subsection{Details on the mechanization}
\renewcommandx{\framep}[2][1,2]{\prim{frame}[][#1][#2][][]}
\renewcommandx{\layer}[2][1,2]{\prim{layer}[][#1][#2][][]}
\renewcommandx{\filler}[2][1,2]{\prim{filler}[][#1][#2][][]}
\renewcommandx{\restrf}[4][1,2,3,4]{\restr{frame}[][#3][#4][#1][#2][][]}
\renewcommandx{\restrl}[4][1,2,3,4]{\restr{layer}[][#3][#4][#1][#2][][]}
\renewcommandx{\restrc}[4][1,2,3,4]{\restr{filler}[][#3][#4][#1][#2][][]}
\renewcommandx{\cohf}[1][1]{\coh{frame}[][][][][][#1][][]}
\renewcommandx{\cohl}[1][1]{\coh{layer}[][][][][][#1][][]}
\renewcommandx{\cohc}[1][1]{\coh{filler}[][][][][][#1][][]}

>>>>

\title{A formalization of semi-simplicial and semi-cubical sets}
\author{
  \textcolor{gray80}{Hugo Herbelin} \\
  \itshape \textcolor{gray80}{Université Paris-Cité, Inria, CNRS, IRIF, Paris}
  \and
  \textcolor{gray80}{Ramkumar Ramachandra} \\
  \itshape \textcolor{gray80}{Université Paris-Cité, Inria, CNRS, IRIF, Paris}
}
\date{}

@@

\begin{document}
\maketitle
\begin{abstract}
  We present a parametricity-based construction of semi-simplicial sets and semi-cubical sets in indexed form, formulated in dependent type theory. The accompanying artifact is an axiom-less~\footnote{Technically speaking, we use some axioms related to functional extensionality, but the requirement arises only from our generalization to $\nu$-types; if $\nu$ is finite (say, unit type corresponding to augmented semi-simplicial types), this requirement vanishes.} Coq formalization that uses sophisticated rewriting techniques and pushes the boundaries of proof assistant technology.
\end{abstract}
\tableofcontents
\newpage

@@

The outline of the paper is as follows. First, we briefly describe semi-simplicial and semi-cubical sets, and our generalization to $\nu$-sets. We then proceed to explain our construction, with the help of diagrams. Before spelling out the details of our formalization in type-theoretic language, we have a section on reasoning in sets. We end with some of the finer details of our mechanization.

Our precise mechanization in Coq is available at \href{https://github.com/artagnon/bonak}{github.com/artagnon/bonak}.

@@

An equivalent definition of UIP is given by \textsf{UIP\_refl}:

@@

% Restriction for frame, layer, and filler
% Drop the universe letter
\renewcommandx{\restrf}[7][1,2,3,4,5,6,7]{\restr{frame}[][#2][#3][#4][#5][#6][#7]}
\renewcommandx{\restrl}[7][1,2,3,4,5,6,7]{\restr{layer}[][#2][#3][#4][#5][#6][#7]}
\renewcommandx{\restrc}[7][1,2,3,4,5,6,7]{\restr{filler}[][#2][#3][#4][#5][#6][#7]}

@@

We now present the construction in \emph{extensional} type theory; i.e. in a type theory with the following reflection rule, where $=$ is propositional equality in some type and $\equiv$ is definitional equality:

@@

Most components of the construction takes inequality constraints as parameters, and we have left implicit that they are satisfied in the table. Comparison of inequality proofs depends on the definition of inequality, which we leave implicit, and assume that we can pick a definition for which inequality proofs are definitionally equal.

@@

Since we have formalized the proof in the Coq proof assistant, which works in intensional type theory, we now present the tables in \emph{intensional} type theory, leaving implicit some details, which we will explain here, and other details, which we will explain in \ref{sec:mech}.

@@

\subsection{Details on the mechanization\label{sec:mech}}
\renewcommandx{\framep}[2][1,2]{\prim{frame}[][#1][#2][][]}
\renewcommandx{\layer}[2][1,2]{\prim{layer}[][#1][#2][][]}
\renewcommandx{\filler}[2][1,2]{\prim{filler}[][#1][#2][][]}
\renewcommandx{\restrf}[4][1,2,3,4]{\restr{frame}[][#3][#4][#1][#2][][]}
\renewcommandx{\restrl}[4][1,2,3,4]{\restr{layer}[][#3][#4][#1][#2][][]}
\renewcommandx{\restrc}[4][1,2,3,4]{\restr{filler}[][#3][#4][#1][#2][][]}
\renewcommandx{\cohf}[1][1]{\coh{frame}[][][][][][#1][][]}
\renewcommandx{\cohl}[1][1]{\coh{layer}[][][][][][#1][][]}
\renewcommandx{\cohc}[1][1]{\coh{filler}[][][][][][#1][][]}
